Subsidiary Recursion in Coq

### 2022-01-31 paper/paper.tex
tweaks to related work

Our work contributes to the program proposed by Owens and Slind, of
broadening the scope of functional programs that can be accommodated
in ITPs~\cite{owens+08}.  The goal of terminating recursion has been
advocated in the literature on programming languages under the name
\emph{strong functional programming}~\cite{turner95}.  Uustalu and
Vene developed a categorical view of a recursion scheme allowing one
level of subsidiary recursion, and illustrated it in Haskell with an
artificial example~\cite{uustalu11}.  In contrast, our scheme allows
arbitrary finite nestings of recursion, and we illustrate it in Coq
with realistic examples.  It seems that generalizing the carriers of
algebras to functors is the critical step enabling such examples. In
our Coq development, we show that Uustalu and Vene's scheme is
derivable from subsidiary recursion. Our method is also similar to the
technique of sized types, in providing a type-based method for
termination~\cite{barthe04}.  With sized types, datatypes are indexed
with abstract sizes, which must then be propagated through code, using
dependent types.  Our \verb|Subrec| does not require dependent types,
resulting in just polymorphically typed code for our examples
(\verb|Subreci|, for proofs, of course does use dependent types).

>>>>

Our work contributes to the program proposed by Owens and Slind, of
broadening the scope of functional programs that can be accommodated
in ITPs~\cite{owens+08}.  The goal of terminating recursion has been
advocated in the literature on programming languages under the name
\emph{strong functional programming}~\cite{turner95}.  Our method is
similar to the technique of sized types, in providing a type-based
method for termination~\cite{barthe04}.  With sized types, datatypes
are indexed with abstract sizes, which must then be propagated through
code, using dependent types.  In contrast, our approach relies just on
polymorphism, and does not require dependent types for writing
subsidiary recursions.  (\verb|Subreci|, for reasoning about such
recursions, of course does use dependent types).

Uustalu and Vene developed a categorical view of a recursion scheme
allowing one level of subsidiary recursion, and illustrated it in
Haskell with an artificial example~\cite{uustalu11}.  In contrast, our
scheme allows arbitrary finite nestings of recursion, and we
illustrate it in Coq with realistic examples.  It seems that
generalizing the carriers of algebras to functors is the critical step
enabling such examples. In our Coq development, we show that Uustalu
and Vene's scheme is derivable from subsidiary recursion.
